---
title: "QBS 181 Final Project"
author: "Chichi Illoh"
date: "2023-11-05"
output: pdf_document
---
```{r}

library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(zoo)
```


```{r}
tesla_deaths = read.csv("/Users/Student/Desktop/QBS 181 Final Project/CleanTesla_Deaths  (1).csv")
tesla_stock = read.csv("/Users/Student/Desktop/QBS 181 Final Project/tesla_stock.csv")
lucid_stock = read.csv("/Users/Student/Desktop/QBS 181 Final Project/Lucid_stock.csv")
head(tesla_deaths)
head(tesla_stock)
head(lucid_stock)
```
# Tesla Stock Dataset Cleaning
```{r}
# Remove the $ in the dataset
tesla_stock$Close.Last <- gsub("\\$", "", tesla_stock$Close.Last)
tesla_stock$Open <- gsub("\\$", "", tesla_stock$Open)
tesla_stock$High <- gsub("\\$", "", tesla_stock$High)
tesla_stock$Low <- gsub("\\$", "", tesla_stock$Low)
tesla_stock$Close <- tesla_stock$Close.Last
tesla_stock$Close.Last <- NULL
```


```{r}
# make prices for Tesla numeric
tesla_stock$Open = as.numeric(tesla_stock$Open)
tesla_stock$High = as.numeric(tesla_stock$High)
tesla_stock$Low = as.numeric(tesla_stock$Low)
tesla_stock$Close = as.numeric(tesla_stock$Close)
```

# Tesla Stock Dataset Cleaning
```{r}
# Remove the $ in the dataset
lucid_stock$Close.Last <- gsub("\\$", "", lucid_stock$Close.Last)
lucid_stock$Open <- gsub("\\$", "", lucid_stock$Open)
lucid_stock$High <- gsub("\\$", "", lucid_stock$High)
lucid_stock$Low <- gsub("\\$", "", lucid_stock$Low)
lucid_stock$Close <- lucid_stock$Close.Last
lucid_stock$Close.Last <- NULL
```


```{r}
# make prices for Lucid numeric
lucid_stock$Open = as.numeric(lucid_stock$Open)
lucid_stock$High = as.numeric(lucid_stock$High)
lucid_stock$Low = as.numeric(lucid_stock$Low)
lucid_stock$Close = as.numeric(lucid_stock$Close)

```
## TidyR: Combining Tesla Stock and Fatal Tesla Car Accidents 

```{r}
tslastock_deaths =tesla_stock %>%
  #select(-origin, -dest) %>% 
  left_join(tesla_deaths, by = "Date")
```

## TidyR: Combining Tesla Stock,Lucid stock, and Fatal Tesla Car Accidents 

```{r}

combined_stocks_death = tslastock_deaths %>%
  inner_join(lucid_stock, by = "Date")
  
```

In Excel NAs are put in as "#N/A". In R this becomes a string and so, each "#N/A" value needs to be converted to a real NA value in R. 
```{r}
na.strings<-c("","#N/A") #set #N/A to be NAs
na_ix<-Reduce("|",lapply(na.strings,"==",combined_stocks_death)) #find the NAs
is.na(combined_stocks_death)<-na_ix #set NAs to the right places
```

Columns are renamed so that Tesla stock and Lucid stock info are differentiated.
```{r}
combined_stocks_death = combined_stocks_death %>% 
  rename("Tesla_Close" = Close.x) %>% 
  rename("Tesla_Volume" = Volume.x) %>%
  rename("Tesla_Open" = Open.x) %>% 
  rename("Tesla_High" = High.x) %>%
  rename("Tesla_Low" = Low.x) %>%
  rename("Lucid_Close" = Close.y) %>% 
  rename("Lucid_Volume" = Volume.y) %>%
  rename("Lucid_Open" = Open.y) %>% 
  rename("Lucid_High" = High.y) %>%
  rename("Lucid_Low" = Low.y) 
```

Convert the Dates in the dataset into a Date object
```{r}
combined_stocks_death = combined_stocks_death%>%
  mutate(Date = mdy(Date))
```


## TidyR: Combining Tesla & Lucid Stock 

Convert the Dates in the dataset into a Date object
```{r}
tesla_stock = tesla_stock%>%
  mutate(Date = mdy(Date))
```

Once again, convert the Dates in the dataset into a Date object
```{r}
lucid_stock = lucid_stock%>%
  mutate(Date = mdy(Date))
```

```{r}
combine_stocks <- tesla_stock%>%
  left_join(lucid_stock, by = "Date")
```


Columns are renamed so that Tesla stock and Lucid stock info are differentiated.
```{r}
library(dplyr)
combine_stocks = combine_stocks %>% 
  rename("Tesla_Close" = Close.x) %>% 
  rename("Tesla_Volume" = Volume.x) %>%
  rename("Tesla_Open" = Open.x) %>% 
  rename("Tesla_High" = High.x) %>%
  rename("Tesla_Low" = Low.x) %>%
  rename("Lucid_Close" = Close.y) %>% 
  rename("Lucid_Volume" = Volume.y) %>%
  rename("Lucid_Open" = Open.y) %>% 
  rename("Lucid_High" = High.y) %>%
  rename("Lucid_Low" = Low.y) 
```


Create csv files of the combined datasets
```{r}
write.csv(combined_stocks_death,file='/Users/Student/Desktop/QBS 181 Final Project/combined_stocks_death.csv')
```

```{r}
write.csv(tslastock_deaths,file='/Users/Student/Desktop/QBS 181 Final Project/tslastock_deaths.csv')
```

```{r}
#write.csv(combine_stocks,file='/Users/admin/new_file.csv')
write.csv(combine_stocks,file='/Users/Student/Desktop/QBS 181 Final Project/combine_stocks.csv')
```







## Exploratory analysis of the data


```{r}
dim(tslastock_deaths)
```

```{r}
library(DataExplorer)
introduce(tslastock_deaths)
```

```{r}
na.strings<-c("","#N/A") #set #N/A to be NAs
na_ix<-Reduce("|",lapply(na.strings,"==",tslastock_deaths)) #find the NAs
is.na(tslastock_deaths)<-na_ix #set NAs to the right places
```


Now lets run again:

```{r}
introduce(tslastock_deaths)
```



# Line plot of Tesla Highs in stock
```{r}
ggplot(data=combine_stocks, aes(x=combine_stocks$Date, y=combine_stocks$Tesla_High, group=1)) +
  geom_line()
```

```{r}
# make prices for TEsla numeric in combined stock

combine_stocks$Tesla_Open = as.numeric(combine_stocks$Tesla_Open)
combine_stocks$Tesla_High = as.numeric(combine_stocks$Tesla_High)
combine_stocks$Tesla_Low = as.numeric(combine_stocks$Tesla_Low)
combine_stocks$Tesla_Close = as.numeric(combine_stocks$Tesla_Close)
```


```{r}
summary(tslastock_deaths)
```

# Visual Exploratory Analysis
```{r}
# for combined stocks
plot(combine_stocks)
```

```{r}
# for tslastock_deaths
plot(tslastock_deaths)
```


# Histograms

```{r}
library(ggplot2)
ggplot(combine_stocks,aes(x=`Tesla_Volume`))+ 
  geom_histogram(bins=20,colour="black",fill="#CA3C25")+
  labs(x="Volume of Tesla stock in a Day", y = "Frequency") #another layer!

```

```{r}
ggplot(combine_stocks,aes(x=`Lucid_Volume`))+ 
  geom_histogram(bins=20,colour="black",fill="#6B717E")+
  labs(x="Volume of Lucid stock in a Day", y = "Frequency") #another layer!

```


```{r}
ggplot(combine_stocks,aes(x=`Tesla_High`))+ 
  geom_histogram(bins=20,colour="black",fill="#CA3C25")+
  labs(x="Tesla Highest Stock Price in a Day", y = "Frequency") #another layer!

```


```{r}
ggplot(combine_stocks,aes(x=`Lucid_High`))+ 
  geom_histogram(bins=20,colour="black",fill="#CA3C25")+
  labs(x="Lucid Highest Stock Price in a Day", y = "Frequency") #another layer!

```

```{r}
ggplot(tslastock_deaths,aes(x=`Deaths`))+ 
  geom_histogram(bins=20,colour="black",fill="#CA3C25")+
  labs(x="Distribution of Deaths in a Car Accidents", y = "Frequency") #another layer!

```

```{r}
ggplot(tslastock_deaths,aes(x=`State`, y=`Deaths`))+ 
  geom_boxplot()+
  labs(x="States", y = "Deaths")

```


```{r}
ggplot(tslastock_deaths,aes(x=`Open`, y=`Deaths`))+ 
  geom_point()+
  labs(x="Open", y = "Deaths")

```


```{r}
ggplot(tslastock_deaths,aes(x=`Autopilot.claimed`, y=`Deaths`))+ 
  geom_boxplot()+
  labs(x="Open", y = "Deaths")

```

```{r}
ggplot(data=subset(tslastock_deaths, !is.na(Model)),aes(x=`Model`, y=`Deaths`))+ 
  geom_bar(stat = "identity", fill = "#CA3C25")+
  labs(x="Tesla Models", y = "Fatal Tesla Accidents")

```